A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}

### 2009-04-10 trunk/textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

% deviation
Another detail to be considered is the uniformity of distribution: all servers must have an usage rate as similar as possible so that the distribution is considered fair. To measure this uniformity, the standard deviation, $\sigma$, was calculated for $u(S)$ of all servers for each simulated algorithm. Figure \ref{gra:baldeviation} shows the variation of $\sigma$ in time.

@@

Even with the servers providing the largest variation in the rate of use of its resources, the \ggp\ was the one with the least overall real weight (which includes overhead: $W_{total} + OverHead$) of all proposed algorithms. Its variant \ggpf\ had the least migrations of users, in exchange for an increase in overhead by 48\%, on the average, compared to the original version of the algorithm. Which of these two is better depends on the specific game. In a real-time game , users contantly migrating between servers can introduce delay and hinder the interaction between players and the \ggpf, as shown, presented a total number of migrations 22\% less than \ggp. Furthermore, it may not be possible to use some kind of ``graceful degradation'' to reduce the quality of the game, needing to save up the servers' resources. In this case, the \ggp\ would be the best option.

>>>>

% deviation
Another detail to be considered is the uniformity of distribution: all servers must have an usage rate as similar as possible so that the distribution is considered fair. To measure this uniformity, the standard deviation, $\sigma$, was calculated for $u(S)$ of all servers for each simulated algorithm. Figure \ref{gra:baldeviation} shows the variation of $\sigma{}(u(S))$ over time.

@@

To summarize the evaluated algorithms, we have: \ggp, which introduced an overhead that increased the system load by 13\% on the average, but it presents the second highest number of migrations of players, and the least fair load distribution; \ggpk, whose overhead increases the system load by 28\%, but it presents the third best number of migrations and its distribution is one the most fair; \ggpf, whose load increase is of 22\%, with the second lowest number of migrations and is also one of the most fair algorithms; and we also have \bfa, which increased the load by 34\%, with the highest number of migrations and not as fair as \ggpk\ or \ggpf. Finally, Ahmed's algorithm increased the system load by 29\%, on the average, with the least number of migrations and a fairness level varying a lot over time.

Although the servers presented the largest variation in the resources usage rate, the \ggp\ was the one with the least overall real weight (which includes overhead: $W_{total} + OverHead$) of all simulated algorithms. Its variant \ggpf\ had the second lowest number of user migrations, along with the second lowest overhead and a fair load distribution, on the average, when compared to the other algorithms. Which of these two is better depends on the specific game. In a real-time game, users constantly migrating between servers can introduce delay and hinder the interaction between players.  \ggpf, or Ahmed's algorithm, would be more suited for this situation. However, when it is not possible to use some kind of ``graceful degradation'' to reduce the quality of the game and save up the servers' resources, \ggp\ would be the best option.
